---
title: "Data 607 - Week 7 Assignment"
author: "Peter Kowalchuk"
date: "October 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Assignment - Working with XML and JSON in R

```{r message=FALSE}
library(kableExtra)
```


##HTML

```{r message=FALSE}
library(rvest)
library(dplyr)
```

```{r}
htmlFile<-"https://raw.githubusercontent.com/pkowalchuk/Data607_Week7_Assignment/master/books.html"
htmlBooks<-tbl_df(as.data.frame(read_html(htmlFile) %>% html_table(header = NA, trim=TRUE, fill=TRUE)))
htmlBooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")
```

##XML

```{r message=FALSE}
library(XML)
library(RCurl)
```


```{r}
xmlfile<-getURL("https://raw.githubusercontent.com/pkowalchuk/Data607_Week7_Assignment/master/books.xml")
xmlBooks<-tbl_df(xmlToDataFrame(xmlfile))
xmlBooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")
```


##JSON

```{r message=FALSE}
library(RJSONIO)
library(jsonlite)
```


```{r}
jsonFile<-"https://raw.githubusercontent.com/pkowalchuk/Data607_Week7_Assignment/master/books.json"
isValidJSON(jsonFile)
jsonbooks<-tbl_df(as.data.frame(fromJSON(jsonFile)))
jsonbooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")

```

##Are the three data frames identical?

Although all three data frames contain the same information, they are not structured the exact same way. All three data tables holde the child data for Authors and Genres a little differently, but we can manipulate the data frames such that the result is the same.

In HTML the different authors and genres are stored in the table as item separated by comas. The same way, this data comes into the data frame as a string where the authors are separated by comas. We could if necesary build a vector with each item holding an author or a genre. This might make working with the data in the data frame easier if operations with individual authors were rewuired.

```{r message=FALSE}
library(stringr)
```

```{r}
htmlBooks$Authors[1]<-str_split(htmlBooks$Authors[1],", ")
htmlBooks$Genres[1]<-str_split(htmlBooks$Genres[1],", ")
htmlBooks$Genres[2]<-str_split(htmlBooks$Genres[2],", ")
htmlBooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")
```

In XML all the data is stored as factors. Althought we might choose to work with the data in this format, we can unfactor the data to make it match what we have in HTML.

```{r message=FALSE}
library(varhandle)
```

```{r}
xmlBooks$Title<-unfactor(xmlBooks$Title)
xmlBooks$Authors<-unfactor(xmlBooks$Authors)
xmlBooks$Genres<-unfactor(xmlBooks$Genres)
xmlBooks$Pages<-as.integer(unfactor(xmlBooks$Pages))
xmlBooks$AvailableInAudio<-unfactor(xmlBooks$AvailableInAudio)
#we split the authors to have then in a vector
xmlBooks$Authors<-c(str_split(xmlBooks$Authors[1],","),xmlBooks$Authors[1],xmlBooks$Authors[2])
#we do the same for genres
xmlBooks$Genres<-c(str_split(xmlBooks$Genres[1],","),str_split(xmlBooks$Genres[2],","),xmlBooks$Genres[3])
xmlBooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")
```


In JSON we already have the Author and Genres columns of the data frame holding vectors for the different data items, so no further manipulation is required. For holding data this format seems the most efficient, with less overhead than XML, and geared towards storing and/or transmitting data rather than presenting it as in HTML. In JSON, we could change the names of the columns to match what we have in the other data frames. We can also change the Pages column to integers.

```{r}
colnames(jsonbooks)<-c('Tittle','Authors','Genres','Pages','AvailableInAudio')
jsonbooks$Pages<-as.integer(jsonbooks$Pages)
jsonbooks %>% kable() %>% kable_styling() %>% scroll_box(width = "910px")
```

Withe these operations we have all three data frames with similar structures

```{r}
htmlBooks
xmlBooks
jsonbooks
```


